### 2022-10-12 paper/main.tex
Update main.tex

\documentclass[sigplan,review,11pt,nonacm,natbib=false]{acmart}
\settopmatter{printfolios=false,printccs=false,printacmref=false}
\usepackage[maxnames=1,minnames=1,maxbibnames=100,natbib=true,citestyle=authoryear,bibstyle=authoryear,doi=false,url=false,isbn=false,isbn=false,backend=biber]{biblatex}
\usepackage{debate}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}
  \renewcommand\ttdefault{cmtt}
\usepackage{csquotes}
\usepackage{ffcode}
\usepackage{CJKutf8}
\usepackage{paralist}
\usepackage{cancel} % to enable \cancel command
\usepackage{anyfontsize} % To get rid of font not found warnings
\usepackage{tabularx} % for special tables
\usepackage{amsmath}
\usepackage{to-be-determined}
\usepackage{href-ul}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tcolorbox} % for algorithm
  \tcbuselibrary{skins}
\usepackage{fdsymbol} % for \mathbb
\usepackage{algpseudocode} % for algorithms
\usepackage{multicol} % for two cols in BNF
\usepackage{pgffor} % to enable \foreach
\usepackage{stmaryrd}
\usepackage{mathtools}
\newtheoremstyle{theorems}
  {} % Space above
  {} % Space below
  {} % Theorem body font % (default is "\upshape")
  {} % Indent amount
  {\scshape} % Theorem head font % (default is \mdseries)
  {.} % Punctuation after theorem head % default: no punctuation
  { } % Space after theorem head
  {} % Theorem head spec
\theoremstyle{theorems}
\newtheorem{eodefinition}{Definition}
\newtheorem{eotheorem}{Theorem}

\newcommand\nospell[1]{#1}
\newcommand\br{\\[-4pt]}
\newcommand\sgm[1]{$\sigma_#1$}
\newcommand\dlt[1]{$\delta_#1$}
\newcommand\m[1]{$\mu_#1$}
\newcommand\figcap[1]{\caption{#1}\Description{#1}}
\newcommand\phic{$\varphi$-calculus}
\newcommand\eo{{\sffamily EO}}
\newcommand\XMIR{{\sffamily XMIR}}
\newcommand\lref[1]{the line no.~\ref{ln:#1}}
\newcommand\lrefs[2]{the lines~\ref{ln:#1}--\ref{ln:#2}}
\newcounter{rule}
\renewcommand\therule{R\arabic{rule}}
\newcommand\rr{\smalltriangleright{}}
\newcommand\rrule[1]{{\scshape\sffamily\ref{rule:#1}}}
\newcommand{\jrule}[1]{%
  \refstepcounter{rule}\label{rule:#1}%
  \text{\textbf{\rrule{#1}}}}
\newenvironment{algo}
  {\newcommand\kw[1]{{\bfseries\sffamily ##1}}
  \newcommand\tab{{\hspace*{1em}}}
  \noindent}
  {}

\addbibresource{refs.bib}

\title{Elimination Of Jump Objects}
\author{author}

\begin{document}

\maketitle

\section{Introduction}
\label{sec:intro}
\input{sections/intro}

\section{Related work}
\label{sec:related}
\input{sections/related}

\section{Definitions and Designations}
\label{sec:defs}
\input{sections/defs}

\section{Sample}
\label{sec:sample}
\input{sections/sample}

\section{Method}
\label{sec:method}
\input{sections/method}

\section{Restrictions}
\label{sec:rest}
\input{sections/restrictions}

\section{Toolkit}
\label{sec:tool}
\input{sections/toolkit}

\section{Discussion}
\label{sec:discus}
\input{sections/discussion}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

\printbibliography

\clearpage

\end{document}