### 2022-10-12 paper/sections/restrictions.tex

Based on the concept of strictly object-oriented programming languages like EO, it is not possible to get rid of \ff{forward}/\ff{backward} objects in all cases.
In such languages, some constructs are impossible to be exactly converted into code in functional languages such as Java, C++, Python, and others.

Let's consider the following example of using \ff{goto} object:
\begin{ffcode}
    goto
    [g]
    seq > @
    foo g
    [x] > foo
    x.forward 25
\end{ffcode}
In this case, the object $g$, from which the \ff{forward}/\ff{backward} objects are called, is passed as an argument to another object, and the \ff{forward} object is already will be dataized inside object $foo$.

Similarly, if the \ff{forward}/\ff{backward} objects encapsulate third-party objects whose logic is not known to us, we cannot do the correct jump transformation:
\begin{ffcode}
    goto
    [g]
    foreign_obj > @
    QQ.io.stdout
    "We don't know the logic\n"
    g.forward TRUE
\end{ffcode}

Thus, we can guarantee transformations of \ff{forward} and \ff{backward} objects only if these objects encapsulate objects whose logic is known to us and these \ff{forward} and \ff{backward} objects are inside the \ff{goto} object.