### 2022-10-12 paper/sections/intro.tex

The problem of eliminating Jump-like statements is known for both computer science and software engineering communities for a few decades and formulated by several scientists, some of which have suggested solutions for this problem.

The presence of jump statements in the source code makes it impossible to use the functional method of analysis.
Other program analysis methods, including those based on symbolic execution techniques, will also benefit from the removal of jump statements, which are well-known obstacles to the precision of analysis, as explained by \citet{baldoni2018survey}.

There are an "object-flow" languages where the execution flow is controlled by object composition.
An example of such a language is EO, which is based on $\varphi$-calculus introduced by \citet{bugayenko2021eolang} and formalized by \citet{kudasov2021}.
This is a strictly object-oriented programming language in which any action is described by creating a new object, or decorating an existing one.

We are trying to answer the following research question: "Is it possible for any program, written in 'object-flow' language like EO and containing GOTO objects, to find a semantically equivalent program, but without GOTO objects?"